PH651 Assignment Five

### 2016-10-09 em/latex/HW2.tex
Add all.

\title{Electromagnetism Assignment One}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

@@

\section{Problem One}
\textbf{Boundaries:}

\begin{align*}
  \phi(\infty) &= 0\\
  \phi(x,0) &= 0\\
  \phi(0,y) &= 0\\
\end{align*}

Guess the solution is four charges in a rectangle. If this solution works, by the Uniqueness Theorem it is the only solution.

\begin{align*}
  \phi &= \phi_1 + \phi_2 + \phi_3 + \phi_4\\
\end{align*}

The potential due to the real point charge is:

\begin{align*}
  \phi_1(x,y) &= \frac{q}{4\pi\epsilon\sqrt{(x-d_1)^2+(y-d_2)^2}}\\
\end{align*}

We let the second and fourth-quadrant points have charge -q, so potential:

\begin{align*}
  \phi_2(x,y) &= \frac{-q}{4\pi\epsilon\sqrt{(x+d_1)^2+(y-d_2)^2}}\\
  \phi_4(x,y) &= \frac{-q}{4\pi\epsilon\sqrt{(x-d_1)^2+(y+d_2)^2}}\\
\end{align*}

and the third-quadrant point have charge +q, so potential:

\begin{align*}
  \phi_3(x,y) &= \frac{q}{4\pi\epsilon\sqrt{(x+d_1)^2+(y+d_2)^2}}\\
\end{align*}

Thus the total potential is given by:

\begin{align*}
  \phi(x,y) &= \frac{q}{4\pi\epsilon}\Big(
  \left((x-d_1)^2+(y-d_2)^2\right)^{-1/2} + \left((x+d_1)^2+(y+d_2)^2\right)^{-1/2}\\
  &-\left((x+d_1)^2+(y-d_2)^2\right)^{-1/2} - \left((x-d_1)^2+(y+d_2)^2\right)^{-1/2}\Big)
\end{align*}

This satisfies our two boundary conditions:

\begin{align*}
  \phi(0,y) &= \frac{q}{4\pi\epsilon}\Big(
  \left(d_1^2+(y-d_2)^2\right)^{-1/2} + \left(d_1^2+(y+d_2)^2\right)^{-1/2}
  -\left(d_1^2+(y-d_2)^2\right)^{-1/2} - \left(d_1^2+(y+d_2)^2\right)^{-1/2}\Big)
  = 0\\
\end{align*}

\begin{align*}
  \phi(x,0) &= \frac{q}{4\pi\epsilon}\Big(
  \left((x-d_1)^2+d_2^2\right)^{-1/2} + \left((x+d_1)^2+d_2^2\right)^{-1/2}
  -\left((x+d_1)^2+d_2^2\right)^{-1/2} - \left((x-d_1)^2+d_2^2\right)^{-1/2}\Big)
  = 0\\
\end{align*}

To find the force on the first-quadrant charge, we use Coulomb's Equation:

%% \begin{align*}
%%   F &= -q\left(\frac{d}{dx} + \frac{d}{dy}\right) \frac{q}{4\pi\epsilon}\Big(
%%   \left((x-d_1)^2+(y-d_2)^2\right)^{-1/2} + \left((x+d_1)^2+(y+d_2)^2\right)^{-1/2}\\
%%   &-\left((x+d_1)^2+(y-d_2)^2\right)^{-1/2} - \left((x-d_1)^2+(y+d_2)^2\right)^{-1/2}\Big)\\
%%   &= \frac{-q^2}{4\pi\epsilon} \Big(
%%   (x-d_1 + y-d_2)\left((x-d_1)^2+(y-d_2)^2\right)^{-3/2} - (x+d_1 + y+d_2)\left((x+d_1)^2+(y+d_2)^2\right)^{-3/2}\\
%%   &+ (x+d_1 + y-d_2)\left((x+d_1)^2+(y-d_2)^2\right)^{-3/2} - (x-d_1 + y+d_2)\left((x-d_1)^2+(y+d_2)^2\right)^{-3/2}\Big)\\
%%   F(d_1,d_2) &= \frac{-q^2}{4\pi\epsilon} \Big(
%%   - (2d_1 + 2d_2)\left(4d_1^2 + 4d_2^2\right)^{-3/2} + 2d_1\left(4d_1^2\right)^{-3/2} - 2d_2\left(4d_2^2\right)^{-3/2}\Big)\\
%%   &= \frac{q^2}{4\pi\epsilon} \left(\frac{(2d_1 + 2d_2)}{\sqrt{\left(4d_1^2 + 4d_2^2\right)}^3} - \frac{1}{4d_1^2} - \frac{1}{4d_2^2}\right)\\
%% \end{align*}

\begin{align*}
  \phi(\vec{r}) &= \frac{q}{\sqrt{2}4\pi\epsilon_0\sqrt{4d_1^2 + 4d_2^2}}\left(\hat{i} + \hat{j}\right) + \frac{-q}{4\pi\epsilon_0\sqrt{4d_1^2}}\hat{i} + \frac{-q}{4\pi\epsilon_0\sqrt{4d_2^2}}\hat{j}\\
  &= \frac{q}{4\pi\epsilon_0}\left(\frac{1}{\sqrt{8d_1^2 + 8d_2^2}} - \frac{1}{\sqrt{4d_1^2}}\right)\hat{i}
  + \frac{q}{4\pi\epsilon_0}\left(\frac{1}{\sqrt{8d_1^2 + 8d_2^2}} - \frac{1}{\sqrt{4d_2^2}}\right)\hat{j}
\end{align*}

If expanded, at large r, $\phi$ should look like a quadrupole scaling with $\frac{1}{r^3}$.\\

\section{Problem Two}
\textbf{Part A}\\
We want to show the following equality is always true:

\begin{align*}
  \int\rho_1\phi_2d^3\vec{r} &= \int\rho_2\phi_1d^3\vec{r}
\end{align*}

We first briefly solve the problem in one dimension to show how it is done:

\begin{align*}
  \int\rho_1\phi_2d^3\vec{r} &= \int\rho_2\phi_1d^3\vec{r}\\
  -\epsilon\int \frac{\partial^2\phi_1}{\partial x^2}\phi_2dx &= -\epsilon\int \frac{\partial^2\phi_2}{\partial x^2}\phi_1dx\\
  \int \frac{\partial^2\phi_1}{\partial x^2}\phi_2dx &= \int \frac{\partial^2\phi_2}{\partial x^2}\phi_1dx\\
  \frac{\partial \phi_1}{\partial x} \phi_2 \Big|_{-\infty}^{\infty} -
  \int \frac{\partial \phi_2}{\partial x}\frac{\partial \phi_1}{\partial x}dx&=
  \frac{\partial \phi_2}{\partial x} \phi_1 \Big|_{-\infty}^{\infty} -
  \int \frac{\partial \phi_1}{\partial x}\frac{\partial \phi_2}{\partial x}dx\\
  \int \frac{\partial \phi_2}{\partial x}\frac{\partial \phi_1}{\partial x}dx&=
  \int \frac{\partial \phi_2}{\partial x}\frac{\partial \phi_1}{\partial x}dx
\end{align*}

Now in 3D. We first apply the Laplace Equation:

\begin{align*}
  -\epsilon_0\int\nabla^2\phi_1\phi_2d^3\vec{r} &= -\epsilon_0\int\nabla^2\phi_2\phi_1d^3\vec{r}
\end{align*}

We then expand this equation into its cartesian representation:

\begin{align*}
  \int\int\int\left(\sum_{n=x,y,z}\frac{\partial^2}{\partial n^2} (\phi_1) \phi_2\right) dxdydz &=
  \int\int\int\left(\sum_{n=x,y,z}\frac{\partial^2}{\partial n^2} (\phi_2) \phi_1\right) dxdydz
\end{align*}

We then use Integration by Parts:

\begin{align*}
  \int\int\sum_{n=x,y,z}\left(\frac{d}{dn}\phi_1\phi_2\Big|_{-\infty}^{\infty} - \int\frac{\partial}{\partial n}\phi_1\frac{\partial}{\partial n}\phi_2dn \right) dxdydz &=
  \int\int\sum_{n=x,y,z}\left(\frac{d}{dn}\phi_2\phi_1\Big|_{-\infty}^{\infty} - \int\frac{\partial}{\partial n}\phi_2\frac{\partial}{\partial n}\phi_1dn \right) dxdydz
\end{align*}

We define our voltages to be zero at infinity, so the UV term of the integration by parts goes to zero, leaving us with:

\begin{align*}
  \int\int\sum_{n=x,y,z}\left(\int\frac{\partial}{\partial n}\phi_1\frac{\partial}{\partial n}\phi_2dn \right) dxdydz &=
  \int\int\sum_{n=x,y,z}\left(\int\frac{\partial}{\partial n}\phi_2\frac{\partial}{\partial n}\phi_1dn \right) dxdydz
\end{align*}

We see that these expressions are identical since multiplication is commutative, thus the initial statement must be true.
\textbf{Part B}\\

We define the voltage on conductor A as: $V_A = V_{AA} + V_{BA}$ where $V_{BA}$ is the potential due to conductor B. We define $V_{AB}$ as the potential on B due to A. We let $\rho_A$ and $\rho_B$ be the charge distributions on A and B respectively, where $\int\rho_AdV = \int\rho_BdV = Q$. We begin with Green's Reciprocity theorem:

\begin{align*}
  \int V_{AB}(\vec{r})\rho_A(\vec{r})dV &= \int V_{BA}(\vec{r})\rho_B(\vec{r})dV\\
\end{align*}

Because A and B are conductors, we know that their potentials $V_{AB}$ and $V_{BA}$ are constant. Thus we can remove them from the integral:

\begin{align*}
  V_{AB}\int\rho_A(\vec{r})dV &= V_{BA}\int\rho_B(\vec{r})dV\\
  V_{AB}Q &= V_{BA}Q\\
\end{align*}

Thus $V_{AB} = V_{BA}$. From this we can see that $V_A = V_{AA} + V_{AB}$.\\

\section{Problem Three}
The solution to the full sphere in an external field $E_0$ was shown in class to be:

\begin{align*}
  \phi(r,\theta) &= \left(-E_0r + \frac{E_0a^3}{r^2}\right)\cos\theta
\end{align*}

given boundaries $\phi(r=a) = 0$ and $\phi(r\rightarrow\infty) = 0$. We notice that our boundaries for the half sphere problem are the same, plus one addition: $\phi(r<a, z=0) = 0$. The above solution satisfies this condition as well. Then by the Uniqueness Theorem this solution must also be valid for our half sphere.\\

The first term $-E_0r\cos\theta$ is the potential from the external field. The second term $\frac{E_0a^3\cos\theta}{r^2}$ is the term due
to the sphere. This term can be rewritten as the dipole term in the multipole expansion:
\begin{align*}
  \frac{E_0a\cos\theta}{r^2} &= \frac{1}{4\pi\epsilon_0}\frac{1}{r^2} \hat{r} \cdot \vec{p}
\end{align*}

where $\vec{p} = 4\pi\epsilon_0E_0 a^3 \hat{z}$. Thus the half sphere can be exactly expressed as a dipole.\\

>>>>

\title{PH651 Assignment Five}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\newcommand{\bra}[1]{\big<#1\big|}
\newcommand{\ket}[1]{\big|#1\big>}
\newcommand{\braket}[2]{\big<#1\big|#2\big>}

@@

\section{Problem One}
To find the bound charges on this cross section, we use the following equations:

\begin{align*}
  \rho_b &= -\nabla \vec{P} = 0\hspace{2cm} \mbox{P is constant}\\
  \sigma_b &= \vec{P} \cdot \vec{n}\\
\end{align*}

The top two faces each make a $60^\circ$ angle with $\vec{P}$, and the bottom one is antiparallel. Thus we see:

\begin{align*}
  \sigma_{b,top} &= \vec{P} \cdot \hat{n} = P\cos\left(60^\circ\right) = \frac P2\\
  \sigma_{b,bot} &= \vec{P} \cdot \hat{n} = P\cos\left(180^\circ\right) = -P\\
\end{align*}

To find total charge we just sum these densities over the surface area of the prism. We notice that there are two top faces and one
bottom face, thus the two surface charges cancel out and the net charge on the object is zero. We expect this for a dielectric with
no free charge.
